---
title: "Comparing `lpdidcsa` and `did` Packages for Difference-in-Differences Estimation"
author: "Olivier Godechot"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
  pdf_document: default
vignette: "%\\VignetteIndexEntry{Comparing lpdidcsa and did Packages}%\\VignetteEngine{knitr::rmarkdown}%\\VignetteEncoding{UTF-8}"
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(error = TRUE, message = FALSE, warning = FALSE)
```

# Introduction

This vignette shows how to use the `lpdidcsa` package for difference-in-differences (DiD) estimation and compares its results with those from the `did` package by Callaway and Sant'Anna. We focus on two estimation methods: inverse probability weighting (IPW) and adjusted regression. Additionally, we explore how `lpdidcsa` handles missing observations, showcasing its partial rebalancing approach.

# Setup

First, load the required packages and set the working directory.

```{r lib}
library(data.table)
library(did)
library(lpdidcsa)
setwd(dirname(rstudioapi::getActiveDocumentContext()$path))
```

# Data Preparation

We use the `mpdta_r` dataset from the `lpdidcsa` package for our analysis.

```{r data}
data(mpdta_r, package = "lpdidcsa")

mpdta_w  <- lpdidcsa_data(mpdta_r, unit = "countyreal", time = "year",
                     dependent = "lemp", treat = "treat")

```

# Comparison of Estimation Methods

## Inverse Probability Weighting (IPW)

### Using `did` Package

```{r csa-ipw}
csa_ipw_det <- att_gt(
  yname = "lemp",
  tname = "year",
  idname = "countyreal",
  gname = "first.treat",
  base_period = "universal",
  control_group = c("notyettreated"),
  est_method = "ipw",
  bstrap = FALSE,
  cband = FALSE,
  xformla = ~lpop,
  data = mpdta_r
)
# summary(csa_ipw_det)
csa_ipw <- aggte(csa_ipw_det, type = "dynamic")
# summary(csa_ipw)
ggdid(csa_ipw)
```

### Using `lpdidcsa` Package

To obtain equivalent results to those of the `did` package, we interact covariates with time dummy variables.

```{r ipw-did}
lpdidcsa_ipw <- lpdidcsa(
  data = mpdta_w,
  meth = "lpdid_ipw",
  dtreat = "dtreat",
  time = "year",
  unit = "countyreal",
  dependent = "lemp",
  controls = "lpop*i(year)"
)
lpdidcsa_ipw$plot
```

### Comparison of Results

```{r ipw-comparison}
cbind(
  csa_ipw$att.egt[csa_ipw$att.egt != 0],
  csa_ipw$se.egt[csa_ipw$att.egt != 0]
) - lpdidcsa_ipw$est[lpdidcsa_ipw$est$estimate != 0, c("estimate", "se")]
```

## Adjusted Regression

### Using `did` Package

```{r adj-did}
csa_adj_det <- att_gt(
  yname = "lemp",
  tname = "year",
  idname = "countyreal",
  gname = "first.treat",
  base_period = "universal",
  control_group = c("notyettreated"),
  est_method = "reg",
  bstrap = FALSE,
  cband = FALSE,
  xformla = ~lpop,
  data = mpdta
)
# summary(csa_adj_det)
csa_adj <- aggte(csa_adj_det, type = "dynamic")
# summary(csa_adj)
ggdid(csa_adj)
```

### Using `lpdidcsa` Package

Here again, to obtain equivalent results to those of the `did` package, we interact covariates with time dummy variables.

```{r adj-lpdidcsa}
lpdidcsa_adj <- lpdidcsa(
  data = mpdta_w,
  meth = "lpdid_adj",
  dtreat = "dtreat",
  time = "year",
  unit = "countyreal",
  dependent = "lemp",
  controls = "lpop*i(year)",
  type_horizon = "wide"
)
lpdidcsa_adj$plot
```

### Comparison of Results

```{r adj-comparison}
cbind(
  csa_adj$att.egt[csa_adj$att.egt != 0],
  csa_adj$se.egt[csa_adj$att.egt != 0]
) - lpdidcsa_adj$est[lpdidcsa_adj$est$estimate != 0, c("estimate", "se")]
```

# Handling Missing Observations

## Creating a Dataset with Missing Observations

```{r missing-data}
set.seed(444)
mpdta_r[, noise := runif(nrow(mpdta_r))]
mpdta_r[, drop := 1 - (noise > 0.1)]
mpdta_na <- mpdta_r[drop == 0, ]
# nrow(mpdta_na)
mpdta_na[, rebal := (sum(!is.na(year)) == 5) * 1, by = countyreal]
# mpdta_na[,.N, by = rebal]
```

## Estimating with Unbalanced Data

### Using `did` Package

```{r unbalanced-did}
csa_unbal_det <- att_gt(
  yname = "lemp",
  tname = "year",
  idname = "countyreal",
  gname = "first.treat",
  base_period = "universal",
  allow_unbalanced_panel = TRUE,
  control_group = c("notyettreated"),
  bstrap = FALSE,
  cband = FALSE,
  data = mpdta_na
)
# summary(csa_unbal_det)
csa_unbal <- aggte(csa_unbal_det, type = "dynamic")
# summary(csa_unbal)
ggdid(csa_unbal)
```

### Using `lpdidcsa` Package with Partial Rebalancing

```{r partial-rebalancing}
mpdta_na_w <- lpdidcsa_data(
  data = mpdta_na,
  treat = "treat",
  dependent = "lemp",
  unit = "countyreal",
  type_horizon = "wide"
)
didpartbal <- lpdidcsa(
  data = mpdta_na_w,
  meth = "lpdid_rw",
  dtreat = "dtreat",
  time = "year",
  unit = "countyreal",
  dependent = "lemp",
  type_horizon = "wide"
)
didpartbal$plot
```

## Comparison of Results with Missing Observations

```{r missing-comparison}
# Difference between CSA's unbalanced estimate and lpdidcsa partially rebalanced estimate
cbind(
  csa_unbal$att.egt[csa_unbal$att.egt != 0],
  csa_unbal$se.egt[csa_unbal$att.egt != 0]
) - didpartbal$est[didpartbal$est$estimate != 0, c("estimate", "se")]

# Difference between CSA's rebalanced estimate and lpdidcsa partially rebalanced estimate
csa_rebal_det <- att_gt(
  yname = "lemp",
  tname = "year",
  idname = "countyreal",
  gname = "first.treat",
  base_period = "universal",
  allow_unbalanced_panel = FALSE,
  control_group = c("notyettreated"),
  bstrap = FALSE,
  cband = FALSE,
  data = mpdta_na
)
csa_rebal <- aggte(csa_rebal_det, type = "dynamic")
cbind(
  csa_rebal$att.egt[csa_rebal$att.egt != 0],
  csa_rebal$se.egt[csa_rebal$att.egt != 0]
) - didpartbal$est[didpartbal$est$estimate != 0, c("estimate", "se")]
```

# Conclusion

This vignette demonstrates the use of the `lpdidcsa` package for DiD estimation and compares its results with the `did` package. The `lpdidcsa` package offers a robust approach to handling missing observations through partial rebalancing, which can yield estimates closer to the "true" values without significant loss of power.

For further details, refer to the package documentation and the original papers by Callaway and Sant'Anna, and the methodology behind `lpdidcsa`.

\`\`\`
